### 2013-04-24 doc/icml2013/theory.tex
Improved bounds for lemma 2 and some proof adjustment

\begin{theorem}[Convergence of Spectral Experts]
\label{thm:convergence}
Each dataset $\sD_p$ (for $p = 1, 2, 3$) consists of $n$ i.i.d.\ points drawn from a mixture
of linear regressions model with parameter $\theta^*$
(having three independent copies simplifies the analysis).
Let $\|x\|_2 \le R$ with probability $1$
and let $\|\beta_h^*\|_2 \le L$ for all $h \in [k]$.
Let $\Sigma_p \eqdef \E[\cvec(x\tp{p})\tp{2}]$, 
and assume $\Sigma_p \succ 0$ for each $p \in \{1,2,3\}$.
Assume $d \ge k$.
Suppose the number of samples is
$n = \max(n_1,n_2)$
where $n_1 = \Omega\left(d^6 R^{12} (\max_{p \in [3]} \sigmamin(\Sigma_p)^{-2}) \log(d) \log(1/\delta)\right)$ and
$n_2 = \Omega\left( \sigma^6 L^6 R^{12} k (\max_{p \in [3]} \sigmamin(\Sigma_p)^{-1}) \frac{\log^3 (1/\delta)}{\epsilon^2} \right)$.
If the regularization strengths $\lambda_n^{(2)}$ and $\lambda_n^{(3)}$ are
set to $\Omega\left(\sigma^3 L^3 R^6 \sqrt{\frac{\log(1/\delta)}{n}}\right)$,
then the parameter estimates $\hat\theta = (\hat\pi, \hat B)$ returned by
\algorithmref{algo:spectral-experts} (with the columns appropriate permuted)
satisfies 
$\|\hat\pi - \pi^*\|_{\infty} \le \epsilon$
and for all $h \in [k]$,
$\|\hat\beta_h - \beta^*_h\|_2 \le \frac{\epsilon}{\sqrt{\pi_h^*}}$.
\end{theorem}

@@

\begin{lemma}[lower bound on restricted strong convexity]
\label{lem:lowRankLower}
Let $\Sigma_p \eqdef \E[\cvec(x\tp{p})\tp{2}]$.
If $$n = \Omega \left( p ~ d^{2p} \log(d) \frac{R^{4p} \log(1/\delta)}{\sigmamin(\Sigma_p)^2} \right),$$
with probability at least $1-\delta$,
$$\kappa(\opX_p) \ge \frac{\sigmamin(\Sigma_p)}{4}.$$
\end{lemma}
\begin{proof}
Expanding the definition of the observation operator:
$\frac{1}{2n} \|\opX_p(\Delta)\|_2^2
= \frac{1}{2n} \sum_{(x,y) \in \sD} \innerp{\Delta}{x\tp{p}}^2$.
Unfolding the tensor and letting $\hat\Sigma_p \eqdef \frac{1}{n} \sum_{(x,y) \in \sD} \cvec(x\tp{p})\tp{2}$,
we have 
$\frac{1}{2n} \|\opX_p(\Delta)\|_2^2
= \frac{1}{2} \trace(\cvec(\Delta)\tp{2} \hat\Sigma_p)
\ge \frac{1}{2} \|\Delta\|_F^2 \sigmamin(\hat\Sigma_p)$.
It remains to show that $\sigmamin(\hat\Sigma_p)$ and $\sigmamin(\Sigma_p)$ are close.
To do this, we first apply Hoeffding's inequality elementwise.
Since $\|x\|_2 \le R$, we have that for each element $(i,j)$,
$|(\hat\Sigma_p)_{ij} - (\Sigma_p)_{ij}| = O(R^{2p}\sqrt{\frac{\log (1/\delta)}{n}})$.
Applying the union bound over the $d^{2p}$ elements of $\hat\Sigma_p - \Sigma_p$,
we have that the max norm is bounded:
$\|\hat\Sigma_p - \Sigma_p\|_\text{\rm max} = O(R^{2p} \sqrt{\frac{p \log(d) \log (1/\delta)}{n}})$.
The max norm times $d^p$ upper bounds the Frobenius norm, which upper bounds the operator norm, so we have that
$\|\hat\Sigma_p - \Sigma_p\|_\text{\rm op} = O(d^p R^{2p} \sqrt{\frac{p \log(d) \log (1/\delta)}{n}})$.
Using the fact that $\sigmamin(\hat\Sigma_p) \ge \sigmamin(\Sigma_p) - \|\hat\Sigma_p - \Sigma_p\|_\text{\rm op}$
yields the result.
\end{proof}

@@

\begin{lemma}[upper bound on adjoint noise]
\label{lem:lowRankUpper}
Let $\opX_p$ be the linear operator previously defined. Then with
probability at least $1-\delta$,
$$\frac1{n} \|\opX_p^*(\eta_p)\|_\op \le O\left(\sigma^3 L^3 R^6 \sqrt{\frac{\log^3(1/\delta)}{n}}\right).$$
for each $p \in \{1,2,3\}$.
\end{lemma}
\begin{proof}
Let $\hat\E_p[f(x,\epsilon,h)]$ denote the empirical expectation over
the examples in dataset $\sD_p$ (recall the $\sD_p$'s are independent to
simplify the analysis).  By definition,
$\frac1n \|\opX_p^*(\eta_p)\|_\op = \hat\E_p[\eta_p(x) x\tp{p}]$
for $p \in \{1,2,3\}$ defined in \refeqn{y1}, \refeqn{y2}, and
\refeqn{y3}. Each $\eta_p(x)$ is composed of several zero-mean random
variables, and since $\|A + B\|_\op \le \|A\|_\op + \|B\|_\op$, it
suffices to consider each term in turn. For each term, we will
condition on $x$ and use standard concentration inequalities to bound
the noise with probability atleast $1-\delta_1$,
{\scriptsize
\begin{align*}
  \| \hat\E_1[ \eta_1(x) x ] \|_{op} &\le 
            \underbrace{ \|\hat\E_1[\innerp{M_1 - \beta_h}{x} x]\|_\op  }_{ O\left( L R^2     \sqrt{\frac{\log(1/\delta_1)}{n}} \right) } +
            \underbrace{ \|\hat\E_1[\epsilon x]\|_\op                   }_{ O\left( R \sigma  \sqrt{\frac{\log(1/\delta_1)}{n}} \right) } \\
  %  &= O\left( \sigma L R^2 \sqrt{\frac{\log(1/\delta_1)}{n}} \right) \\
    \| \hat\E_2[ \eta_2(x) x\tp{2} ] \|_{op} &\le 
            \underbrace{ \|\hat\E_2[\innerp{\beta_h\tp{2} - M_2}{x\tp{2}} x\tp{2}]\|_\op }_{ O\left( L^2 R^4      \sqrt{\frac{\log^2(1/\delta_1)}{n}} \right) } \\
   &\quad + \underbrace{ \|\hat\E_2[2 \epsilon \innerp{\beta_h}{x} x\tp{2}]\|_\op        }_{ O\left( \sigma L R^3 \sqrt{\frac{\log  (1/\delta_1)}{n}} \right) } 
          + \underbrace{ \|\hat\E_2[(\epsilon^2 - \sigma^2) x\tp{2}]\|_\op               }_{ O\left( \sigma^2 R^2 \sqrt{\frac{\log^2(1/\delta_1)}{n}} \right) } \\
  %&= O\left( \sigma^2 L^2 R^4 \sqrt{\frac{\log^2(1/\delta_1)}{n}} \right) \\
  \| \hat\E_3[ \eta_3(x) x\tp{3} ] \|_{op} &\le 
            \underbrace{ \|\hat\E_3[\innerp{\beta_h\tp{3} - M_3}{x\tp{3}}                      x\tp{3}]\|_\op }_{O\left( L^3 R^6        \sqrt{\frac{\log  (1/\delta_1)}{n}} \right) } \\
   &\quad + \underbrace{ \|\hat\E_3[3 \epsilon \innerp{\beta_h\tp{2}}{x\tp{2}}                 x\tp{3}]\|_\op }_{O\left( \sigma L^2 R^5 \sqrt{\frac{\log^2(1/\delta_1)}{n}} \right) } 
          + \underbrace{ \|\hat\E_3[\epsilon^3                                  x\tp{3}]\|_\op }_{O\left( \sigma^3 R^3   \sqrt{\frac{\log  (1/\delta_1)}{n}} \right) } \\
   &\quad + \underbrace{ \|\hat\E_3[3(\epsilon^2 \innerp{\beta_h}{x} -\sigma^2 \innerp{\hat M_1}{x}) x\tp{3}]\|_\op }_{O\left( (\sigma^3 L R^4 + \sigma^3 R^4) \sqrt{\frac{\log^3(1/\delta_1)}{n}} \right) } \\
  %&= O\left( \sigma^3 L^3 R^6 \sqrt{\frac{\log^3(1/\delta_1)}{n}} \right).
\end{align*}
}
There are two additional considerations when bounding the terms in $\eta_3$.
First, to bound $\hat\E_3[\epsilon^3]$, we use Lemma 19 of \cite{hsu13spherical}.
Second, there is an additional bias since the estimator uses $\hat M_1$ instead of $M_1$; this bias is small by the previous bound (and is constructed on an independent dataset, $\sD_1$).
Therefore, we incur another $O(\sigma^2 \cdot R \sigma \sqrt{\frac{\log(1/\delta_1)}{n}})$.
We can handle the other terms using conventional means, the details of which can be found in \todo{Appendix B}.

%First, consider $\eta_1(x)$, which contains two terms.
%The first term is $\innerp{M_1 - \beta_h}{x}$.
%We have that $\|M_1 - \beta_h\|_2 \le 2L$,
%so by Hoeffding's inequality, and using the fact that $\|x\|_2 \le R$,
%with probability $1-\delta_1$,
%$\|\hat\E_1[\innerp{M_1 - \beta_h}{x} x]\|_\op = O(R^2 L \sqrt{\frac{\log(1/\delta_1)}{n}})$.
%The second term is $\epsilon \sim \normal{0}{\sigma^2}$,
%with probability $1-\delta_1$,
%$\|\hat\E_1[\epsilon x]\|_\op = O(R \sigma \sqrt{\frac{\log(1/\delta_1)}{n}})$.
%
%Now, consider $\eta_2(x)$, which contains three terms.
%Using similar techniques, we have that
%$\|\hat\E_2[\eta_2(x) x\tp{2}]\|_\op = O(R^2 (L^2 R^2 + \sigma L R + \sigma^2) \sqrt{\frac{\log(1/\delta_1)}{n}})$.

% Finally, $\eta_3(x)$ contains five terms plus an additional bias since the estimator uses $\hat M_1$ rather than $M_1$.
% There are two additional considerations.
% First, to bound $\hat\E_3[\epsilon^3]$, we use Lemma 19 of \cite{hsu13spherical}.
% Second, the additional bias is small by the previous bound (and it is constructed on $\sD_1$, which is independent of $\sD_3$
% used to construct $\hat M_3$.
% Therefore, we incur another $O(\sigma^2 \cdot R \sigma \sqrt{\frac{\log(1/\delta_1)}{n}})$.
% We can handle the other terms using conventional means.
% Combining everything, we obtain
% $\|\hat\E_3[\eta_3(x) x\tp{3}]\|_\op = O(R^3 (L^3 R^3 + \sigma L^2 R^2 + \sigma^2 L R + \sigma^3 + \sigma^3 R) \sqrt{\frac{\log^3(1/\delta_1)}{n}})$.

>>>>

\begin{theorem}[Convergence of Spectral Experts]
\label{thm:convergence}
Each dataset $\sD_p$ (for $p = 1, 2, 3$) consists of $n$ i.i.d.\ points drawn from a mixture
of linear regressions model with parameter $\theta^*$
(having three independent copies simplifies the analysis).
Let $\|x\|_2 \le R$ with probability $1$
and let $\|\beta_h^*\|_2 \le L$ for all $h \in [k]$.
Let $\Sigma_p \eqdef \E[\cvec(x\tp{p})\tp{2}]$, 
and assume $\Sigma_p \succ 0$ for each $p \in \{1,2,3\}$.
Assume $d \ge k$.
Suppose the number of samples is
$n = \max(n_1,n_2)$
where $n_1 = \Omega\left(R^{12} (\max_{p \in [3]} \sigmamin(\Sigma_p)^{-2}) \log(1/\delta)\right)$ and
$n_2 = \Omega\left( \sigma^6 L^6 R^{12} k (\max_{p \in [3]} \sigmamin(\Sigma_p)^{-1}) \frac{\log^3 (1/\delta)}{\epsilon^2} \right)$.
If the regularization strengths $\lambda_n^{(2)}$ and $\lambda_n^{(3)}$ are
set to $\Omega\left(\sigma^3 L^3 R^6 \sqrt{\frac{\log(1/\delta)}{n}}\right)$,
then the parameter estimates $\hat\theta = (\hat\pi, \hat B)$ returned by
\algorithmref{algo:spectral-experts} (with the columns appropriate permuted)
satisfies 
$\|\hat\pi - \pi^*\|_{\infty} \le \epsilon$
and for all $h \in [k]$,
$\|\hat\beta_h - \beta^*_h\|_2 \le \frac{\epsilon}{\sqrt{\pi_h^*}}$.
\end{theorem}

@@

\begin{lemma}[lower bound on restricted strong convexity]
\label{lem:lowRankLower}
Let $\Sigma_p \eqdef \E[\cvec(x\tp{p})\tp{2}]$.
If $$n = \Omega \left( \frac{R^{4p} \log(1/\delta)}{\sigmamin(\Sigma_p)^2} \right),$$
with probability at least $1-\delta$,
$$\kappa(\opX_p) \ge \frac{\sigmamin(\Sigma_p)}{4}.$$
\end{lemma}
\begin{proof}
Expanding the definition of the observation operator:
$\frac{1}{2n} \|\opX_p(\Delta)\|_2^2
= \frac{1}{2n} \sum_{(x,y) \in \sD} \innerp{\Delta}{x\tp{p}}^2$.
Unfolding the tensor and letting $\hat\Sigma_p \eqdef \frac{1}{n} \sum_{(x,y) \in \sD} \cvec(x\tp{p})\tp{2}$,
we have 
$\frac{1}{2n} \|\opX_p(\Delta)\|_2^2
= \frac{1}{2} \trace(\cvec(\Delta)\tp{2} \hat\Sigma_p)
\ge \frac{1}{2} \|\Delta\|_F^2 \sigmamin(\hat\Sigma_p)$.
By Weyl's theorem, $\sigmamin(\hat\Sigma_p) \ge \sigmamin(\Sigma_p) - \|\hat\Sigma_p - \Sigma_p\|_{op}$. 
Thus, it remains to show that $\|\hat\Sigma_p - \Sigma_p\|_{op} \le \|\hat\Sigma_p - \Sigma_p\|_{F}$ is small. 
To do so, we apply a matrix concentration result for the Frobenius norm\citationneeded; with probability at least $1 - \delta$, 
$\| \hat\Sigma_p - \Sigma_p \|_F = O\left( M \sqrt{\frac{\log(1/\delta)}{n}} \right)$,
where $M$ is an upper bound on $\| \Sigma_p \|_F$. Since $\|x\|_2 \le R$, we can use the bound $\| \Sigma_p \|_F \le R^{2p}$ in the above to get the stated result.

%To do this, we first apply Hoeffding's inequality elementwise.
%Since $\|x\|_2 \le R$, we have that for each element $(i,j)$,
%$|(\hat\Sigma_p)_{ij} - (\Sigma_p)_{ij}| = O(R^{2p}\sqrt{\frac{\log (1/\delta)}{n}})$.
%Applying the union bound over the $d^{2p}$ elements of $\hat\Sigma_p - \Sigma_p$,
%we have that the max norm is bounded:
%$\|\hat\Sigma_p - \Sigma_p\|_\text{\rm max} = O(R^{2p} \sqrt{\frac{p \log(d) \log (1/\delta)}{n}})$.
%The max norm times $d^p$ upper bounds the Frobenius norm, which upper bounds the operator norm, so we have that
%$\|\hat\Sigma_p - \Sigma_p\|_\text{\rm op} = O(d^p R^{2p} \sqrt{\frac{p \log(d) \log (1/\delta)}{n}})$.
%Using the fact that $\sigmamin(\hat\Sigma_p) \ge \sigmamin(\Sigma_p) - \|\hat\Sigma_p - \Sigma_p\|_\text{\rm op}$
%yields the result.
\end{proof}

@@

\begin{lemma}[upper bound on adjoint noise]
\label{lem:lowRankUpper}
Let $\opX_p$ be the linear operator previously defined. Then with
probability at least $1-\delta$,
$$\frac1{n} \|\opX_p^*(\eta_p)\|_\op \le O\left(\sigma^3 L^3 R^6 \sqrt{\frac{\log^3(1/\delta)}{n}}\right).$$
for each $p \in \{1,2,3\}$.
\end{lemma}
\begin{proof}
Let $\hat\E_p[f(x,\epsilon,h)]$ denote the empirical expectation over
the examples in dataset $\sD_p$ (recall the $\sD_p$'s are independent to
simplify the analysis).  By definition,
$\frac1n \|\opX_p^*(\eta_p)\|_\op = \hat\E_p[\eta_p(x) x\tp{p}]$
for $p \in \{1,2,3\}$ defined in \refeqn{y1}, \refeqn{y2}, and
\refeqn{y3}. Each $\eta_p(x)$ is composed of several zero-mean random
variables, and since $\|A + B\|_\op \le \|A\|_\op + \|B\|_\op$, it
suffices to consider each term in turn. 
For each term, we will
condition on $x$ and use standard concentration inequalities to bound
the noise.

For example, consider $\eta_1(x)$, which has two terms. 
The first term is $\innerp{M_1 - \beta_h}{x}$.  We have that $\|M_1 - \beta_h\|_2 \le 2L$, so by Hoeffding's inequality, and using the fact that $\|x\|_2 \le R$, with probability $1-\delta_1$, $\|\hat\E_1[\innerp{M_1 - \beta_h}{x} x]\|_2 = O(R^2 L \sqrt{\frac{\log(1/\delta_1)}{n}})$. 
The second term is $\epsilon \sim \normal{0}{\sigma^2}$, thus with probability $1-\delta_1$, $\|\hat\E_1[\epsilon x]\|_2 = O(R \sigma \sqrt{\frac{\log(1/\delta_1)}{n}})$. 
The bounds on the terms presented below hold with probability atleast $1-\delta_1$ and are proved similarly.
{\scriptsize
\begin{align*}
%   \| \hat\E_1[ \eta_1(x) x ] \|_{op} &\le 
%             \underbrace{ \|\hat\E_1[\innerp{M_1 - \beta_h}{x} x]\|_\op  }_{ O\left( L R^2     \sqrt{\frac{\log(1/\delta_1)}{n}} \right) } +
%             \underbrace{ \|\hat\E_1[\epsilon x]\|_\op                   }_{ O\left( R \sigma  \sqrt{\frac{\log(1/\delta_1)}{n}} \right) } \\
  %  &= O\left( \sigma L R^2 \sqrt{\frac{\log(1/\delta_1)}{n}} \right) \\
    \| \hat\E_2[ \eta_2(x) x\tp{2} ] \|_{op} &\le 
            \underbrace{ \|\hat\E_2[\innerp{\beta_h\tp{2} - M_2}{x\tp{2}} x\tp{2}]\|_\op }_{ O\left( L^2 R^4      \sqrt{\frac{\log^2(1/\delta_1)}{n}} \right) } \\
   &\quad + \underbrace{ \|\hat\E_2[2 \epsilon \innerp{\beta_h}{x} x\tp{2}]\|_\op        }_{ O\left( \sigma L R^3 \sqrt{\frac{\log  (1/\delta_1)}{n}} \right) } 
          + \underbrace{ \|\hat\E_2[(\epsilon^2 - \sigma^2) x\tp{2}]\|_\op               }_{ O\left( \sigma^2 R^2 \sqrt{\frac{\log^2(1/\delta_1)}{n}} \right) } \\
  %&= O\left( \sigma^2 L^2 R^4 \sqrt{\frac{\log^2(1/\delta_1)}{n}} \right) \\
  \| \hat\E_3[ \eta_3(x) x\tp{3} ] \|_{op} &\le 
            \underbrace{ \|\hat\E_3[\innerp{\beta_h\tp{3} - M_3}{x\tp{3}}                      x\tp{3}]\|_\op }_{O\left( L^3 R^6        \sqrt{\frac{\log  (1/\delta_1)}{n}} \right) } \\
   &\quad + \underbrace{ \|\hat\E_3[3 \epsilon \innerp{\beta_h\tp{2}}{x\tp{2}}                 x\tp{3}]\|_\op }_{O\left( \sigma L^2 R^5 \sqrt{\frac{\log^2(1/\delta_1)}{n}} \right) } 
          + \underbrace{ \|\hat\E_3[\epsilon^3                                  x\tp{3}]\|_\op }_{O\left( \sigma^3 R^3   \sqrt{\frac{\log  (1/\delta_1)}{n}} \right) } \\
   &\quad + \underbrace{ \|\hat\E_3[3(\epsilon^2 \innerp{\beta_h}{x} -\sigma^2 \innerp{\hat M_1}{x}) x\tp{3}]\|_\op }_{O\left( (\sigma^3 L R^4 + \sigma^3 R^4) \sqrt{\frac{\log^3(1/\delta_1)}{n}} \right) } \\
  %&= O\left( \sigma^3 L^3 R^6 \sqrt{\frac{\log^3(1/\delta_1)}{n}} \right).
\end{align*}
}
There are two additional considerations when bounding the terms in $\eta_3$.
First, to bound $\hat\E_3[\epsilon^3]$, we use Lemma 19 of \cite{hsu13spherical}.
Second, there is an additional bias since the estimator uses $\hat M_1$ instead of $M_1$; this bias is small by the previous bound (and is constructed on an independent dataset, $\sD_1$).
Therefore, we incur another $O(\sigma^2 \cdot R \sigma \sqrt{\frac{\log(1/\delta_1)}{n}})$.
%We can handle the other terms using conventional means, the details of which can be found in \appendixref{sec:proofs}.
